---
title: 'Section 3: Structure analyses'
output:
  html_document:
    df_print: paged
---

*Note: the analyses in this notebook expect certain files to have already been created by the preprocessing.Rmd and structure.ipynb notebooks*

```{r message=FALSE,results='hide'}
library(tidyverse)
library(lme4)
library(ggthemes)
library(tm)
library(tidyboot)
library(xtable)
library(entropy)
library(useful)
library(viridis)
library(broom)
library(broom.mixed)

library(ggthemes)
library(cowplot)

source('../utils/analysis_helpers.R')
```

# 3 | RESULTS: CHARACTERIZING THE DYNAMICS OF STRUCTURE

## Section 3.1: The effect of matcher feedback on reduction

```{r warning=FALSE,message=FALSE,results='hide'}
# Import uncollapsed version of data to examine matcher
message_df = read_csv('../../data/tangramsSequential.csv') %>%
  mutate(numRawWords = str_count(contents, "\\S+")) 

meta_message_df = message_df %>% 
  group_by(gameid, trialNum, repetitionNum, intendedName) %>% 
  summarize(correct = mean(correct), numRawWords = sum(numRawWords))

listenerMsgs <- message_df %>%
  group_by(gameid, repetitionNum, role, intendedName) %>% #taskVersion
  summarize(individualM = n()) %>%
  ungroup() %>%
  complete(gameid, intendedName, repetitionNum, role, fill = list(individualM = 0)) %>%
  spread(role, individualM)  %>% 
  left_join(meta_message_df)

listenerReduction <- listenerMsgs %>%
   group_by(repetitionNum, gameid) %>%
   summarize(matcherTalks = sum(matcher) > 0) %>%
   group_by(repetitionNum) %>%
   tidyboot_mean(matcherTalks) %>%
   mutate(measure = '# listener messages')  
```

Statistics for reduction in matcher message rates

```{r}
listenerMsgs %>% 
  mutate(matcherBin = matcher > 0) %>%
  glmer(matcherBin ~ repetitionNum + 
          (1 + repetitionNum | gameid) + 
          (1  | intendedName), 
        data = ., 
        family = 'binomial') %>%
  tidy()
```

### Visualization of matcher rates (Fig. 2B)

```{r}
ggplot(listenerReduction, aes(x = repetitionNum, y = mean*100)) +
  geom_ribbon(aes(ymax = 100 * ci_upper, 
                  ymin = 100 * ci_lower), 
              alpha = 0.1, color = F) +
  geom_line() +
  ylab('% games where \n matcher sent message') +
  xlab('repetition #') +
  ylim(0,100) +
  scale_x_continuous(breaks = round(seq(1, 6, by = 1),1)) +
  theme_cowplot() +
  theme(aspect.ratio = 1)

ggsave('../../writing/figs/listenerResponses.pdf', 
       height = 8, width = 12, units = 'cm', useDingbats = F)
```

### mini-analysis of what is actually contained in listener response messages

```{r}
# Count question marks
message_df %>% 
  filter(role == 'matcher') %>% 
  mutate(containQ = str_detect(contents, "\\?")) %>% 
  group_by(containQ) %>% 
  tally() %>% 
  ungroup() %>% 
  mutate(proportion = n / sum(n))
```

### Were directors sensitive to ground-truth accuracy feedback?

```{r}
# Extract reduction in response to feedback for 
# both repetition and trial-by-trial baseline
feedback <- listenerMsgs %>% 
  mutate(matcherBin = matcher > 0) %>%
  group_by(gameid) %>% arrange(gameid, trialNum) %>%  
  mutate(trial_prevcorrect = lag(correct),
         trial_reduction = log(numRawWords/lag(numRawWords))) %>%
  group_by(gameid, intendedName) %>% arrange(gameid, repetitionNum) %>% 
  mutate(rep_prevcorrect = lag(correct), 
         rep_reduction = log(numRawWords/lag(numRawWords))) %>% 
  gather(comparison, val, trial_prevcorrect : rep_reduction) %>%
  separate(comparison, into = c('comparison', 'metric')) %>%
  spread(metric, val) %>%
  group_by(gameid) %>%
  # only look at games where at least one error was made 
  # (otherwise no rounds after error to compare)
  filter(sum(prevcorrect == 0, na.rm = T) > 0) %>%
  ungroup() %>%
  filter(!is.na(prevcorrect)) %>%
  filter(!is.na(reduction)) %>%
  mutate(prevcorrectFactor = factor(ifelse(prevcorrect == 1, 'correct', 'incorrect')),
         comparisonNumeric = ifelse(comparison == "rep", 1, 0))
```


```{r}
feedback %>% 
  filter(repetitionNum == 1) %>% 
  group_by(matcherBin) %>% 
  summarize(m = 1 - mean(correct))

tab <- table((feedback %>% filter(repetitionNum == 1))$matcherBin, 
             (feedback %>% filter(repetitionNum == 1))$correct)

chisq.test(tab)
```

```{r}
feedback %>%
  filter(comparison == 'rep') %>%
  lmer(reduction ~ 1 + prevcorrect + I(repetitionNum) + 
         (1 + prevcorrect + I(repetitionNum) | gameid), 
       data = .,
      control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=100000))) %>%
  tidy()
```

Test interaction with traditional regression b/c even minimal lmer does not converge

```{r}
lm(reduction ~ prevcorrectFactor * comparison, data = feedback) %>%
  tidy()
```

Backup Bayesian mixed-effects model mentioned in footnote

```{r warning=FALSE,message=FALSE, cache=TRUE}
library(brms)
library(rstan)
brm(reduction ~ prevcorrectFactor * comparison + 
               (1 + prevcorrectFactor * comparison | gameid),
    chains = 1,
    iter = 5000,
    data = feedback) %>%
  tidy()
```

Make Fig. 2C

```{r}
feedback.toplot <- feedback %>% 
  filter(comparison == 'rep') %>%
  group_by(repetitionNum, prevcorrect) %>%
  tidyboot_mean(column = reduction, na.rm =T, nboot = 1000) 

feedback.toplot %>%
  ungroup() %>%
  ggplot(aes(x = repetitionNum, y = empirical_stat, color = factor(prevcorrect))) +
    geom_point() +
    geom_line() +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), 
                  width = 0, position = 'dodge') +
    geom_hline(aes(yintercept = 0)) +
    theme_cowplot()  +
    theme(aspect.ratio = 1) +
    ylab('(log) ratio of # words vs. previous round') +
    ylim(-0.8, 0.8)

ggsave('../../writing/figs/responseToFeedback.pdf', 
       height = 8, width = 16, units = 'cm', useDingbats = F)
```

## Section 3.2 | Breaking down the structure of reduction

```{r, echo=FALSE,message=FALSE}
tagged_df <- read_csv('../outputs/posTagged_tangramsSequential_collapsed.csv')
```

### Pure reduction in utterance length (Fig. 2A)

```{r}
to.plot <- tagged_df %>%
   group_by(repetitionNum) %>% 
   tidyboot_mean(numWords, na.rm = T, nboot = 500) 

to.plot %>%
   mutate(measure = '# words per tangram') %>%
   ggplot(aes(x = repetitionNum, y = empirical_stat)) +
    geom_ribbon(aes(ymax = ci_upper, ymin = ci_lower), alpha= 0.2) +
    geom_line() +
    xlab('repetition #') +
    ylab('mean # words used \n per tangram') +
    scale_x_continuous(breaks = round(seq(1, 6, by = 1),1)) +
    ylim(c(0, 20)) +
    theme_cowplot()  + 
    theme(aspect.ratio = 1)

ggsave('../../writing/figs/reduction.pdf',
        height = 8, width = 12, units = 'cm', useDingbats = F)
```

Statistics:

```{r}
tagged_df %>% 
   group_by(gameid, trialNum, repetitionNum, intendedName) %>% 
   summarize(numRawWords = sum(numWords)) %>%
  lmer(log(numRawWords) ~ poly(repetitionNum,2) + 
               (1 + repetitionNum | gameid) + 
               (1 + repetitionNum | intendedName), 
       data = .,
       control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=100000))) %>%
  tidy()
```

Supplemental check -- we examined the number of total words a director produced for a tangram, but at the beginning when there is more back-and-forth dialogue, this set of words also includes responses to the matcher. Does the number of words in the *initial* description go down over time, too? (i.e. the description produced before the listener chimes in).

```{r}
message_df %>%
  group_by(gameid, trialNum) %>%
  mutate(uttNumInTrial = row_number()) %>%
  filter(role == 'matcher') %>%
  summarize(firstListenerUtt = first(uttNumInTrial)) %>%
  right_join(message_df) %>%
  mutate(firstListenerUtt = ifelse(is.na(firstListenerUtt), 1000, firstListenerUtt)) %>%
  group_by(gameid, trialNum, repetitionNum) %>%
  mutate(uttNumInTrial = row_number()) %>%
  filter(uttNumInTrial < firstListenerUtt) %>%
  summarize(m = sum(numRawWords)) %>%
  group_by(repetitionNum) %>%
  summarize(m = mean(m))
```

### Table 1

What are the unigrams/bigrams/trigrams that drop the most in frequency across whole data set?

```{r}
makeNGramTable(15)
```

### Reduction in parts of speech (Fig. 3)

examine reduction in different parts of speech

```{r message=F}
read_csv('../outputs/posTagged_tangramsSequential_collapsed.csv') %>%
  group_by(repetitionNum) %>%
  summarize(numWords = sum(numWords),
            nouns = sum(NOUNcount)/sum(numWords),
            verbs = sum(VERBcount)/sum(numWords),
            dets= sum(DETcount)/sum(numWords),
            preps = sum(ADPcount)/sum(numWords),
            conjunctions_pronouns = (sum(CCONJcount) + sum(SCONJcount) +
                                       sum(PRONcount))/sum(numWords),
            adjectives = sum(ADJcount)/sum(numWords)) %>%
  mutate(OTHER = (1 - nouns - verbs - dets -conjunctions_pronouns -
                      preps - adjectives)) %>%
  gather(POS, prop, nouns:OTHER) %>%
  mutate(POS = factor(POS, levels = c('nouns',  'verbs', 'preps', 'adjectives',
                               'conjunctions_pronouns', 'dets', 'OTHER'))) %>%
  select(repetitionNum, POS, prop) %>%
  ggplot(aes(x = repetitionNum, y = prop, fill = POS)) +
    geom_area(alpha=0.6 , size=1, colour="black") +
    scale_fill_brewer(palette = "Set1") +
    ggthemes::theme_few() +
    scale_x_continuous(breaks = c(1,2,3,4,5,6))+
    theme(aspect.ratio = 2.5) +
    ylab('% words')
ggsave('../../writing/figs/posChange.pdf',
        height = 8, width = 16, units = 'cm', useDingbats = F)
```

Explicitly look at reduction for each POS.

```{r message=FALSE}
posReduction <- tagged_df %>%
  group_by(gameid, repetitionNum) %>%
  summarize(numWords = sum(numWords),
            numMessages = length(gameid),
            nouns = sum(NOUNcount),
            verbs = sum(VERBcount),
            dets= sum(DETcount),
            pronouns = sum(PRONcount),
            preps = sum(ADPcount),
            adverbs = sum(ADVcount),
            conjunctions = sum(CCONJcount) + sum(SCONJcount),
            adjectives = sum(ADJcount)) %>%
  gather(POS, count, nouns:adjectives) %>%
  select(gameid, repetitionNum, POS, count) %>%
  # Need to put in ids to spread
  rowwise() %>%
  mutate(id = row_number(),
         repetitionNum = paste0('rep', repetitionNum, collapse = '')) %>%
  spread(repetitionNum, count) %>%
  # Compute % reduction from first to last round
  mutate(diffPct = (rep1 - rep6)/rep1) %>%
  group_by(POS) %>%
  # Filter out handful of people who skipped first round w/ negative %...
  filter(diffPct >= 0) %>%
  filter(POS != "OTHER") %>%
  mutate(cat = case_when(POS %in% c('dets', 'pronouns', 'preps', 'conjunctions') ~ 'closed', 
                         POS == 'adverbs' ~ '?',
                         TRUE ~ 'open'),
         cat = factor(cat, levels = c('closed', '?', 'open'))) %>%
  # rearrange
  transform(POS=reorder(POS, -diffPct))

posReduction.boot <- posReduction %>%
  group_by(POS, cat) %>%
  # Take mean & se over participants
  tidyboot_mean(column = diffPct, na.rm = T)

ggplot(posReduction.boot, aes(x = POS, y = mean, fill = cat)) +
  geom_bar(stat = 'identity') +
  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = .1)+
  ylab("absolute % reduction") +
  xlab("Part of Speech category")  +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_fill_manual(values=c('tomato3', 'gray', 'orange')) +
  theme_few() +
  coord_flip() +
  theme(aspect.ratio = 1)

ggsave('../../writing/figs/posReduction.pdf',
        height = 8, width = 16, units = 'cm', useDingbats = F)
```

```{r}
detReductionRate <- (posReduction.boot %>% filter(POS == 'dets'))$mean * 100
conjReductionRate <- (posReduction.boot %>% filter(POS == 'conjunctions'))$mean * 100
pronounReductionRate <- (posReduction.boot %>% filter(POS == 'pronouns'))$mean * 100
nounReductionRate <- (posReduction.boot %>% filter(POS == 'nouns'))$mean * 100

cat('determiners reduce by', detReductionRate, '\n')
cat('conjunctions reduce by', conjReductionRate, '\n')
cat('pronouns reduce by', pronounReductionRate, '\n')
cat('nouns reduce by', nounReductionRate, '\n')
```

Stats for open class vs. closed class

```{r}
library(broom.mixed)
library(lmerTest)
posReduction %>% 
  filter(cat != '?') %>%
  lmer(diffPct ~ cat + (1 | gameid) + (1 | POS),
       data = .) %>%
  tidy()
```

### Reduction in syntactic units (+ Supplemental Fig. 10)

Get list of all words dropped on each round...

```{r}
syntactic <- read_csv('../outputs/permuted_dependency_distribution.csv') 

syntactic %>%   
  filter(repetitionNum == 'avg') %>%
  group_by(baselineName) %>%
  summarize(lower = min(value), upper = max(value))

syntactic %>% 
  mutate(repetitionNum = fct_rev(factor(paste0('repetition ', repetitionNum)))) %>%
  ggplot(aes(x = repetitionNum, y= value, fill=baselineName, color = baselineName)) +
    geom_point(data = subset(syntactic, baselineName == 'true'), size = 4) +
    geom_violin(data = subset(syntactic, baselineName != 'true'))  +
    theme_few() +
    ylab('mean dependency length between dropped words') +
    xlab('') +
    coord_flip() +
    #scale_x_continuous(breaks = NULL) +
    theme(aspect.ratio = 1/2)

ggsave('../../writing/figs/syntactic_reduction_baselines.pdf',
       height = 8, width = 16, units = 'cm', useDingbats = F)
```

