---
title: "tangramsReference"
output: 
  html_document:
    toc: true
    toc_depth: 2

---

# Import data

```{r}
library(tidyverse)
library(lme4)
library(ggthemes)
library(tm)
library(tidyboot)
library(xtable)
library(ggthemes)
library(entropy)
library(useful)
library(viridis)
library(broom)
library(gganimate)

# reticulate lets us import npy files with numpy python package
library(reticulate)
np <- import("numpy")

source('./utils/analysis_helpers.R')
```

# Dynamics of content

```{r}
d <- read_csv('../data/tangramsSequential_collapsed.csv', col_types = 'ciiccil')
```


## What determines whether a label gets conventionalized?

Note that the only way to get an exact 0 is if word never appeared with tangram... so we want to filter these out ()

```{r}
distinctiveness_d <- read.csv("outputs/PMI.csv", header = TRUE) %>%
  mutate(finalRoundMatch = ifelse(finalRoundMatch == 'True', 1, 0)) %>%
  filter(!is.na(tf.idf)) %>%
  mutate(num_tangrams_occurred_with =  round(1 / ((2^tf.idf) /12))) 
```

look at how often final round occurred on early round
```{r}
distinctiveness_d %>% mutate(didNotOccurOnFirstRound = tf.idf == 0) %>% 
  group_by(finalRoundMatch, didNotOccurOnFirstRound) %>% tally()
```

stats

```{r}
summary(glmer(finalRoundMatch ~ num_tangrams_occurred_with + (1 + num_tangrams_occurred_with | intendedName) + (1 + num_tangrams_occurred_with | gameid), 
              family = 'binomial', control = glmerControl(optCtrl = list(maxfun = 200000)), data = distinctiveness_d %>%   filter(tf.idf != 0 )))
```

Not exactly sure how to look at this... 

```{r}
distinctiveness_d %>%
  filter(tf.idf != 0) %>%
  group_by(num_tangrams_occurred_with) %>%
  tidyboot_mean(finalRoundMatch) %>%
  rename(sample_size = n) %>%
  ggplot(aes(x = as.integer(num_tangrams_occurred_with), y =empirical_stat)) +
    geom_point(aes(size = sample_size), stat = 'identity') +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.1) +
    geom_smooth(method = 'lm', formula = y ~ poly(x, 2), se = F) +
    ylim(0, .22) +
    ylab('% appearing in final utterance') +
    xlab('distinctiveness of word in initial utterance (# tangrams)') +
    xlim(0,12) +
    scale_x_continuous(breaks = c(1,3, 5,7,9,11))+
    theme_few() +
    theme(aspect.ratio = 1)

ggsave('../writing/tangrams/figs/distinctiveness.pdf',  height = 7, width = 10, units = 'cm', useDingbats = F)
```

Alternatively, can do a nonparametric analysis: draw a random word from each tangram/gameid pair and look at the percentage that match with round 6... This gives a null distribution. Then we can take the highest PMI word (or words) for each tangram/gameid pair and look at the percentage of *those* that match. We see that it's much higher than expected under the null.

```{r}
# TODO: get red to show up in legend
nonparametric_ds = map_dbl(seq_len(1000), ~{ distinctiveness_d %>%
  group_by(intendedName, gameid) %>%
  sample_n(1) %>%
    ungroup() %>%
    summarize(m = mean(finalRoundMatch)) %>%
    pull(m)})
    
highest <- map_dbl(seq_len(1000), ~{ distinctiveness_d %>%
  group_by(intendedName, gameid) %>%
  filter(num_tangrams_occurred_with == min(num_tangrams_occurred_with)) %>%
  sample_n(1) %>%
  ungroup() %>%
  summarize(m = mean(finalRoundMatch)) %>%
  pull(m)})
  
cat('range for permuted=', sort(nonparametric_ds)[1], sort(nonparametric_ds)[1000])
cat('range for highest=', sort(highest)[1], sort(highest)[1000])

```

## semantic similarity analyses

```{r}
M_mat = read_csv('./outputs/meta_tangrams_embeddings.csv', col_types = "nccnlc", na = c('[nan]')) %>%
  mutate(feature_ind = X1 + 1, # Have to correct for R's (stupid) 1-indexing...
         gameID = gameid,
         target = intendedName,
         repetition = as.numeric(repetitionNum)) #%>% 
F_mat <- np$load('outputs/feats_tangrams_embeddings_rawavg.npy')
```

### overall btw vs. within similarity

```{r}
combined.df <- compute_within_vs_across(M_mat, F_mat)

true_dprime = dprime(combined.df)
permuted_dprimes <- map_dbl(seq_len(1000), ~ dprime(combined.df %>% mutate(source = sample(source))))
cat('CI for permuted=', sort(permuted_dprimes)[25], sort(permuted_dprimes)[975])
cat('true=', true_dprime)

combined.df %>%
  ggplot(aes(x = empirical_stat, fill = source)) +
    geom_density(adjust = 1.5, alpha = .5) +
    #facet_wrap(~ target) +
    xlab('pairwise cosine similarity') +
    theme_few()

ggsave('../writing/tangrams/figs/across_vs_within.pdf',  height = 7, width = 10, units = 'cm', useDingbats = F)
```

### Baseline across tangrams within game

One concern is that these effects are an artifact of our method of creating utterance vectors by averaging word vectors. Maybe this procedure causes all long utterances to be more similar?

It may look like different pairs are diverging and getting further apart, but expect, however, that at each point in time, 

It appears that similarity is slightly *higher* among utterances for the same tangram across different games than for different tangrams within the same game. 

```{r}
acrossGames <- M_mat %>%
    group_by(target, repetitionNum) %>%
    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = 'cosine'),
                          .$gameid)) %>%
    filter(dim1 != dim2) %>%
  rename(gameid = dim1) %>%
    group_by(repetitionNum,gameid) %>%
  summarize(empirical_stat = mean(sim, na.rm = T)) %>% mutate(source ='across')

diffTangramWithinGames <- M_mat %>%
  group_by(gameid, repetitionNum) %>%
  do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = 'cosine'),
                        .$target)) %>%
  summarize(empirical_stat = mean(sim, na.rm = T)) %>%
  filter(!is.na(empirical_stat)) %>%
  mutate(source = 'across tangrams in game') 

bind_rows(acrossGames,diffTangramWithinGames) %>%
  spread(source, empirical_stat) %>%
  mutate(diff = across - `across tangrams in game`) %>%
  group_by(repetitionNum) %>%
  tidyboot_mean(diff, na.rm = T)
  ggplot(aes(x = as.character(repetitionNum), y = empirical_stat, color = source)) +
    geom_violin()
```


### Drift

Stats 

```{r}
permuted.drift <- compute_permuted_estimates(M_mat, F_mat, 'drift', 100)
cat('CI for drift analysis=', sort(permuted.drift)[2], sort(permuted.drift)[98])

true_lmer.drift <- make_within_df(M_mat, F_mat, 'cosine') %>% filter(rep1 == 1) %>% ungroup()
true_lmer.drift.out <- lmer(sim ~ poly(rep2,2) + (1 + poly(rep2, 2)  | gameID) + (1 + poly(rep2, 2)| target), data = true_lmer.drift)
cat('true for drift analysis=', (tidy(true_lmer.drift.out, effects = 'fixed') %>% filter(term == 'poly(rep2, 2)1'))$estimate)
summary(true_lmer.drift.out)
```

Viz (raw data)

```{r}
true_lmer.drift %>% 
  unite('repdiff', rep1, rep2, sep = '->') %>%
  bind_rows(data.frame(repdiff ='1->1', sim = 1)) %>%
  group_by(repdiff) %>%
  mutate(group_mean = mean(sim, na.rm = T)) %>%
  ggplot(aes(x = repdiff, y = sim)) +
    geom_boxplot() +#width = .1, outlier.size=0
    scale_x_discrete(labels = c('(1,1)','(1,2)', '(1,3)', '(1,4)', '(1,5)' ,'(1,6)','(1,7)')) +
    ylim(0, 1) +
    ylab('cosine similarity') +
    ggtitle('drift from initial within game') +
    theme_few() +
    xlab('repetition pair') +
    guides(color = F, fill = F) +
    theme(aspect.ratio = 1, legend.position = c(0.5, 0.8), text = element_text(size=18), 
          element_line(size=1), element_rect(size=2, color="#00000"))

ggsave('../writing/tangrams/figs/drift_boxplot.pdf', height = 10, width = 10, units = 'cm', useDingbats = F)
```

Viz (mean estimates w/ baseline)

```{r}
empirical_baselines.drift <- combine_empirical_and_baselines(M_mat, F_mat, 'drift', 10) %>%
    bind_rows(data.frame(IV ='1->1', empirical_stat=1, ci_lower=1, 
                         ci_upper=1, sample_id='empirical')) %>%
    bind_rows(compute_across_similarity(M_mat, F_mat, 'baseline', 'cosine') %>%
                filter(IV == 1) %>% select(-n, -mean) %>%
                mutate(IV = '1->1')) 

empirical_baselines.drift %>%
  mutate(sample_id = fct_relevel(sample_id, 'empirical', 'baseline')) %>%
  ggplot(aes(x = IV, y = empirical_stat, color = sample_id, linetype = sample_id, group = sample_id)) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0) + 
    geom_line()  +
    #geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0, data = empirical_baselines.drift %>% filter(sample_id == 'empirical')) +
    #geom_line()  +
    #geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.5, color = NA, data = empirical_baselines.drift %>% filter(sample_id == 'baseline')) +
    scale_x_discrete(labels = c('(1,1)','(1,2)', '(1,3)', '(1,4)', '(1,5)' ,'(1,6)','(1,7)')) +
    ylim(0.4, 1) +
    ylab('cosine similarity') +
    ggtitle('drift from initial within game') +
    theme_few() +
    xlab('repetition pair') +
    guides(color = F, fill = F) +
    theme(aspect.ratio = 1, legend.position = c(0.8, 0.8), text = element_text(size=18), 
          element_line(size=1), element_rect(size=2, color="#00000"))

ggsave('../writing/tangrams/figs/drift_mean.pdf', height = 10, width = 10, units = 'cm', useDingbats = F)

```

### Within-interaction convergence

Stats 

```{r}
true_lmer.within <- make_within_df(M_mat, F_mat, 'cosine') %>% filter(rep2 == rep1 + 1) %>% ungroup()
true_lmer.within.out <- lmer(sim ~ poly(rep1,2) + (1 + poly(rep1,2) | gameID) + (1+ poly(rep1,2)| target), data = true_lmer.within)
summary(true_lmer.within.out)
cat('true for within analysis=', (tidy(true_lmer.within.out, effects = 'fixed') %>% filter(term == 'poly(rep1, 2)1'))$estimate)

permuted.within <- compute_permuted_estimates(M_mat, F_mat, 'within', 100)
cat('CI for within analysis=', sort(permuted.within)[2], sort(permuted.within)[98])
```

Viz (raw data)

```{r}
true_lmer.within %>% 
  unite('repdiff', rep1, rep2, sep = '->') %>%
  group_by(repdiff) %>%
  mutate(group_mean = mean(sim, na.rm = T)) %>%
  ggplot(aes(x = repdiff, y = sim)) +
    geom_boxplot() +#width = .1, outlier.size=0
    scale_x_discrete(labels = c('(1,1)','(1,2)', '(1,3)', '(1,4)', '(1,5)' ,'(1,6)','(1,7)')) +
    ylim(0, 1) +
    ylab('cosine similarity') +
    ggtitle('convergence within game') +
    theme_few() +
    xlab('repetition pair') +
    guides(color = F, fill = F) +
    theme(aspect.ratio = 1, legend.position = c(0.5, 0.8), text = element_text(size=18), 
          element_line(size=1), element_rect(size=2, color="#00000"))
ggsave('../writing/tangrams/figs/stability_boxplot.pdf', height = 10, width = 10, units = 'cm', useDingbats = F)

```

Viz (mean estimates w/ baseline)

```{r}
combined.within <- combine_empirical_and_baselines(M_mat, F_mat, 'within', 2)

combined.within %>%
  mutate(sample_id = fct_relevel(sample_id, 'empirical', 'baseline')) %>%
  ggplot(aes(x = IV, y = empirical_stat, color = sample_id, linetype = sample_id, group = sample_id)) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0) + 
#                  data = combined.within %>% filter(sample_id == 'empirical')) +
    geom_line()  +
    # geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.5, color = NA, 
    #             data = combined.within %>% filter(sample_id == 'baseline')) +
    scale_x_discrete(labels = c('(1,2)','(2,3)', '(3,4)', '(4,5)', '(5,6)' )) +
    ylim(0.4, 1) +
    ylab('cosine similarity') +
    ggtitle('convergence within game') +
    theme_few() +
    xlab('repetition pair') +
    guides(color = F, fill = F) +
    theme(aspect.ratio = 1, legend.position = c(0.6, 0.5), text = element_text(size=18), 
          element_line(size=1), element_rect(size=2, color="#00000"))

ggsave('../writing/tangrams/figs/stability_mean.pdf', height = 10, width = 10, units = 'cm', useDingbats = F)

```

### Between-interaction divergence 

Stats 

```{r}
true_lmer.across <- M_mat %>% 
   group_by(target, repetition) %>%  
   do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = 'cosine'), .$gameID)) %>% 
   #unite(col = 'gamepair', dim1, dim2) %>%
  ungroup() %>%
  mutate(target = factor(target))
true_lmer.across.out <- lmer(sim ~ poly(repetition, 2) + (1 +poly(repetition, 2)  | target), control = lmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')),  data = true_lmer.across,verbose = TRUE)
summary(true_lmer.across.out)
# , 

cat('true for within analysis=', (tidy(true_lmer.across.out, effects = 'fixed') %>% filter(term == 'poly(repetition, 2)1'))$estimate)

permuted.across <- compute_permuted_estimates(M_mat, F_mat, 'across', 100)
cat('CI for within analysis=', sort(permuted.across)[2], sort(permuted.across)[99])
```

Viz (full data)

```{r}
true_lmer.across %>%
  ggplot(aes(x = factor(repetition), y = sim)) +
    geom_boxplot() +#width = .1, outlier.size=0
    scale_x_discrete(labels = c('(1,1)','(1,2)', '(1,3)', '(1,4)', '(1,5)' ,'(1,6)','(1,7)')) +
    ylim(0, 1) +
    ylab('cosine similarity') +
    ggtitle('convergence within game') +
    theme_few() +
    xlab('repetition pair') +
    guides(color = F, fill = F) +
    theme(aspect.ratio = 1, legend.position = c(0.5, 0.8), text = element_text(size=18), 
          element_line(size=1), element_rect(size=2, color="#00000"))
ggsave('../writing/tangrams/figs/divergence_boxplot.pdf', height = 10, width = 10, units = 'cm', useDingbats = F)

```

Viz (aggregated)

```{r}
combined.across <- combine_empirical_and_baselines(M_mat, F_mat, 'across', 100)
combined.across %>%
  mutate(sample_id = fct_relevel(sample_id, 'empirical', 'baseline')) %>%
ggplot(aes(x = IV, y = empirical_stat, 
           fill = sample_id, color = sample_id, linetype = sample_id, group = sample_id)) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0) +
    geom_line()  +
    scale_x_continuous(breaks = c(1,2,3,4,5,6), labels = c(1,2,3,4,5,6)) +
    ylim(.4, 1) +
    ylab('cosine similarity') +
    ggtitle('divergence across pairs') +
    theme_few() +
    xlab('repetition') +
    guides(color = F, fill = F) +
    theme(aspect.ratio = 1, legend.position = c(0.5, 0.8), text = element_text(size=18), 
          element_line(size=1), element_rect(size=2, color="#00000"))
ggsave('../writing/tangrams/figs/divergence_mean.pdf', height = 10, width = 10, units = 'cm', useDingbats = F)

```

## some controls

i'm making this claim in the paper that people start out with something shared in how they talk about particular tangrams; people are closer together at the beginning in a tangram-specific way. but maybe everyone is using long utterances at the beginning which makes it look similar. if that were true, then sampling of the content wouldn't matter. if you scramble words across tangrams within games and recompute the averaged utterance vector, then you'd expect that permutation to be equally similar. instead, we find that similarity drops.

BUT maybe they're saying similar things for every tangram?

2. take actual rep 1 things people say and scramble game ids... (how much is due to tangram differentiation overall vs. pragmatic informativity pressures)


```{r}
M_mat_scrambled <- read_csv('./outputs/meta_tangrams_embeddings_scrambled0.csv', col_types = "nccnlc", na = c('[nan]')) %>%
    mutate(feature_ind = X1 + 1, # Have to correct for R's (stupid) 1-indexing...
         gameID = gameid,
         target = intendedName,
         repetition = as.numeric(repetitionNum)) 

raw_npy_names <- dir(path = './outputs/', pattern = "feats_tangrams_embeddings_rawavg_scrambled[0-9]*", full.names = T)
raw_data <- map(raw_npy_names, ~ {
   M_mat_scrambled %>% 
    group_by(target, repetition) %>%  
    do(flatten_sim_matrix(get_sim_matrix(., np$load(.x), method = 'cosine'), .$gameID)) %>% 
    ungroup() %>%
    group_by(repetition) %>%
    summarize(m = mean(sim, na.rm = T)) %>%
    mutate(sample_id = .x)
  }) %>%
  reduce(rbind)               # read all csvs and combine into single data frame

raw_data %>% group_by(repetition) %>% summarize(min = min(m), max = max(m)) %>% left_join(true_lmer.across %>% group_by(repetition) %>% summarize(empirical = mean(sim, na.rm = T)))
```

## tsne visualizations

```{r}
tsne <- read_csv('outputs/embeddings.csv') %>%
  left_join(read_csv('../data/tangramsSequential_collapsed.csv') ) %>%
  select(gameid, intendedName, repetitionNum, x_tsne, y_tsne, contents, feats_type) %>%
  mutate(r = useful::cart2pol(x_tsne, y_tsne)$r,
         theta = useful::cart2pol(x_tsne, y_tsne)$theta) %>%
  group_by(gameid, feats_type, intendedName) %>%
  arrange(repetitionNum) %>%
  mutate(finalTheta = last(theta)) 

tsne.toplot <- tsne %>% 
  filter(intendedName == 'C') %>% 
  filter(repetitionNum %in% c(1, 6)) %>%
  filter(!is.na(x_tsne)) %>%
  filter(feats_type == 'raw_avg') %>%
  group_by(feats_type, gameid) %>%
  mutate(next_x = lead(x_tsne) ,
         next_y = lead(y_tsne))

ggplot(tsne.toplot, aes(x = x_tsne, y = y_tsne, color = finalTheta)) +
    geom_point(data = subset(tsne.toplot, repetitionNum == 1),
               size = 1) +
    geom_segment(aes(xend = next_x, yend = next_y), arrow.fill = NULL, 
                 arrow = arrow(length=unit(0.30,"cm"), angle=15, type = "closed"), ) +
    # remove text for official version...
    geom_text(aes(label=contents), size = 1)+
    theme_few(20) +
    ggtitle('pca + tsne embeddings') +
    scale_shape_manual(values = c(21)) +
    scale_alpha_continuous(range = c(0.5, 1))+
    scale_color_gradientn(colours = viridis(5))+
      theme(axis.title=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank()) +
    labs(x = "", y = "") +
    guides(color = F, shape = F, alpha = F) +
    theme(aspect.ratio = 1)  #+
# make it big to see text... remove
ggsave(filename = '../writing/tangrams/figs/tsne-tangramC.pdf', width = 15, height = 15)
```

MDS ...

```{r}
ggplot(tsne %>% filter(intendedName == 'C'), 
       aes(x = x_tsne, y = y_tsne, group = gameid, frame = paste0('repetition: ', repetitionNum))) +
  geom_line(alpha = 0.2) +
  geom_point(aes(shape = factor(-repetitionNum), 
                 color = finalTheta, 
                 alpha = factor(repetitionNum)),
             size = 1.5, stroke = 2,
             data = tsne %>% filter(repetitionNum %in% c(1,6))) +
  #facet_wrap(~ intendedName) +
  theme_few(20) +
  ggtitle('mds embeddings') +
  scale_shape_manual(values = c(16, 21)) +
  scale_alpha_discrete(range = c(0.5, 1))+
  scale_color_gradientn(colours = tableau_color_pal()(3))+
  guides(color = F, shape = F, alpha = F) +
  theme(aspect.ratio = 1) 

ggsave('../writing/tangrams/figs/mds_embeddings.pdf', width = 15, height = 15)
```

TODO: how do tangrams cluster?

```{r}

#? 
```

## Discrete analyses: How much is within-game similarity disrupted by scrambling utterances across games?

Calculate entropy under permutation tests

Note that we explicitly do *not* normalize entropies, as this
removes the contribution of a bigger vocabulary (as we would expect)

```{r}
getCounts <- function(contents) {
  corpus <- Corpus(VectorSource(paste(contents, collapse = " ")))
  return(colSums(as.matrix(DocumentTermMatrix(corpus, control = list(
    removePunctuation = TRUE, stopwords = TRUE, stemming = TRUE)))))
}

getCI <- function(vector) {
  l = length(vector)
  upper95 = ceiling(l*0.975)
  lower95 = floor(l*0.025)
  return(data.frame(list(ci_lower = sort(vector)[lower95], 
                         ci_upper = sort(vector)[upper95],
                         empirical_stat = mean(sort(vector)))))
}

permutationTest <- function(d, sample_id) {
  # scramble across games within each repetition/tangram
  permuted = d %>%
    group_by(repetitionNum, intendedName) %>%
    mutate(permutation = sample(contents)) %>%
    group_by(gameid, intendedName) %>%
    summarize(ent = entropy(getCounts(permutation))) %>%
    group_by(intendedName) %>%
    summarize(meanEnt = mean(ent)) %>%
    mutate(sample_id = sample_id)
  return(permuted)
}

permutations <- map_dfr(seq_len(1000), ~permutationTest(d, .x)) %>%
  group_by(intendedName) %>% 
  do(., getCI(.$meanEnt)) %>% mutate(sample_id = 'permuted')

trueEntropy <- d %>%
  group_by(gameid, intendedName) %>%
  summarize(trueEntropy = entropy(getCounts(contents))) %>%
  group_by(intendedName) %>%
  tidyboot_mean(trueEntropy) %>%
  mutate(sample_id = 'empirical') %>% 
  select(intendedName, ci_lower,ci_upper, empirical_stat, sample_id)

discrete.out <- bind_rows(permutations, trueEntropy)
```

Plotted 

```{r}
ggplot(discrete.out, aes(x = empirical_stat, y = fct_reorder(factor(intendedName), empirical_stat), 
                         color = sample_id)) +
  geom_point() +
  geom_errorbarh(aes(xmax = ci_upper, xmin = ci_lower), height = 0) +
  theme_few() +
  ylab('tangram') +
  xlab('mean entropy') +
  #coord_flip() +
  scale_x_continuous(breaks = NULL) +
  theme(aspect.ratio = 1/2)

ggsave('../writing/tangrams/figs/permutedDiscrete.pdf', height = 8, width = 16, units = 'cm', useDingbats = F)
```

```{r}
summary(lmer(ent ~ permutation + (1 + permutation | gameid) + (1 + permutation | intendedName), 
             data = allEntropies))
```

# Supplemental analyses

## Slightly sketchy way of looking at arbitrariness by comparing across-pair and within-pair entropy

```{r}
library(entropy)

withinPair <- oldGrams %>% 
  group_by(gameid, tangramRef) %>%
  summarize(ent = entropy(colSums(as.matrix(DocumentTermMatrix(Corpus(VectorSource(paste(cleanMsg, collapse = " ")))))))) %>%
  group_by(tangramRef) %>%
  summarize(withinEnt = mean(ent), withinSE = sd(ent)/sqrt(length(ent))) 

withinPair

acrossPair <- oldGrams %>% 
  filter(tangramRef != 0) %>%
  filter(tangramRef != 'None') %>%
  filter(tangramRef != '*') %>%
  group_by(tangramRef, repetitionNum) %>% 
  summarize(acrossEnt = entropy(colSums(as.matrix(DocumentTermMatrix(Corpus(VectorSource(paste(cleanMsg, collapse = " ")))))))) %>%
  left_join(withinPair, by = "tangramRef") %>%
  gather(type, entropy, acrossEnt, withinEnt)

ggplot(acrossPair, aes(x = repetitionNum, y = entropy, 
                       color = type, linetype = tangramRef)) +
  geom_line()
```

Or we could look at both on each half? 

```{r}
library(entropy)

withinPair <- oldGrams %>% 
  filter(tangram != 0) %>%
  mutate(half = factor(ifelse(repetitionNum <= 3, "beg", "end"))) %>%
  group_by(gameid, tangram, half) %>%
  summarize(ent = entropy(colSums(as.matrix(DocumentTermMatrix(Corpus(VectorSource(paste(cleanMsg, collapse = " ")))))))) %>%
  group_by(tangram, half) %>%
  summarize(withinEnt = mean(ent), withinSE = sd(ent)/sqrt(length(ent))) 

withinPair

acrossPair <- oldGrams %>% 
  filter(tangram != 0) %>%
  mutate(half = factor(ifelse(repetitionNum <= 3, "beg", "end"))) %>%
  group_by(tangram, half) %>% 
  summarize(acrossEnt = entropy(colSums(as.matrix(DocumentTermMatrix(Corpus(VectorSource(paste(cleanMsg, collapse = " ")))))))) %>%
  left_join(withinPair, by = c("tangram", "half")) %>%
  gather(type, entropy, acrossEnt, withinEnt)

acrossPair

ggplot(acrossPair,
       aes(x = half, y = entropy, 
           color = tangram, linetype = type, group = interaction(tangram, type))) +
  geom_line()
```

## tsne animation

```{r}
p <- ggplot(tsne %>% filter(repetitionNum %in% c(1,6)) %>% group_by(gameid, intendedName) %>% filter(length(gameid) == 2), aes(x = x_tsne, y = y_tsne, group = interaction(gameid,intendedName))) +
  #geom_line(alpha = 0.2) +
  geom_point(aes(#shape = factor(-repetitionNum),
                 color = finalTheta,
                 alpha = 1
    ),
             size = 1.5, stroke = 2
             ) +
  facet_wrap(~ intendedName) +
  theme_few(20) +
  ggtitle('pca + tsne embeddings') +
  scale_shape_manual(values = c(21)) +
  scale_alpha_continuous(range = c(0.5, 1))+
  scale_color_gradientn(colours = viridis(5))+
  guides(color = F, shape = F, alpha = F) +
  theme(aspect.ratio = 1)  +
  labs(title = "Rep. {floor(frame_time)}", x = "", y = "") +
  theme(axis.title=element_blank(),
      axis.text=element_blank(),
      axis.ticks=element_blank()) +
  transition_time(as.numeric(repetitionNum)) # speed it up!

   
animate(p, nframes = 50)
anim_save(filename = '../writing/tangrams/figs/tsne-animation.gif')
```
